Above all, I want to express my gratitude to my supervisors, Prof. Harith Alani and Dr. Alistair Willis, for their unwavering support and guidance throughout my research. They were always available to discuss any problems or inquiries I had and provided valuable feedback.
I would like to give a special thanks to Prof. Harith Alani for his thought-provoking questions that helped me stay on track with my research. His talent for simplifying complex ideas into understandable concepts demonstrated that every problem can be solved with the proper approach and clear explanations.
I am grateful for the technical assistance provided by Dr. Alistair Willis, which enabled me to effectively frame and focus the most intricate section of my thesis. Through his invaluable support, I gained valuable insight into designing strong experiments and critically analysing results.

% guidance showed me the importance of good writing in research. Hopefully, his ability to spot the smallest typos and grammatical errors made me a better writer.

I would like to express my gratitude to my friends and colleagues at the Knowledge Media Institute (KMi), Computing \& Communications (C\&C) and generally in the Open University family, for their support throughout my studies. Their assistance has been invaluable to me. Paula, Angel, Angelo, Lucas, Lara, Alba, Agnese, Gianluca, Perla, Tommy, Pedro, Thais, Patrizia, Julian, Miriam, Cecilia and Paco: you all contributed to making life in Milton Keynes better.
I also express my gratitude to two former KMi members who are the initial cause of my ending up in Milton Keynes: Ilaria and Emanuele. Without you, a lot of things would have been substantially different.

And thanks to the two musical bands that accompanied me during this 3-year period, \emph{KMinstruments} and \emph{I Gabbiani}, that provided a good and motivating distraction and made the time more enjoyable. 

Finally, I would like to thank my mum, dad, brother, and sister for supporting me even in the distance over all these years, and spurring me to continue.% and not give up.

Last but not least, I would like to thank Giorgia.
%for her unofficial supervising abilities and for standing on my side.
%and motivation to push me all the times I lost my energy.
Your support has been an inexhaustible source of comfort and motivation for me in all the tough times. Your positive energy and determined encouragement always kept me going, even when things seemed impossible. Your ability to put things into perspective, and focus on the finish line, helped me remain determined and focused on my goals. I am truly grateful for your persistent support and will always cherish it.

%Her ability to deal with a PhD researcher, stressed and anxious about the experiments failing and this never-ending thesis, cannot be expressed in words.
